%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

%\squo{We have seen that computer programming is an art,
%    because it applies accumulated knowledge to the world,
%    because it requires skill and ingenuity, and especially
%    because it produces objects of beauty.}{Donald Knuth}

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

None of this work could have been completed without tremendous help, guidance, support, and love. I'd like to say a few
words about all the people who came together to realize this thesis.

My partner, Emily, has been my biggest supporter throughout my studies, and has stood by me through good times and bad
during my PhD. She has given me feedback, helped me formulate and make sense of my thoughts and myself, and, with great
patience and love, has helped me along the way through the marathon that is graduate school. Perhaps one day I can repay
her kindness and patience for these graduate school years, but that may take a lifetime.

None of this work would have been possible without the guidance of my dissertation committee, who, in no small way,
provided me the support and environment I needed to succeed. In no particular order, I want to profusely thank:

\begin{itemize}
    \item Peter Alvaro, \emph{whose blue-sky insights have shaped this work and story like no other.}
    \item Ethan Miller, \emph{with his concrete and meticulous approach to systems engineering.}
    \item Darrell Long, \emph{whose kindness is matched equally well with invaluable critical feedback.}
    \item Pankaj Mehra, \emph{whose propensity to probe at the difficult questions of research always leads to better understanding.}
\end{itemize}

The above list is in no way meant to imply that these qualities are completely unique to each person; on the contrary,
I look up to these people precisely because they all share these vital qualities in a mentor. They have all provided me
guidance and training for analyzing problems, building mental models, probing those models, designing experiments,
running and managing them, collecting and processing data, turning that into new knowledge, and of course, communicating
that knowledge. And, perhaps most important of all, zooming out and looking at the big picture. I can only hope to
emulate their brilliance going forward.

I will never forget the unending support from my lab mates, in no particular order, Dev Purandare, Peter Wilcox, Michael Usher,
Achilles Benetopoulos, Esteban Ramos, Matt Bryson, Staunton Sample, Yan Li, Lincoln Thurlow, Oceane Bel, Ken Chang,
Sinjoni Mukhopadhyay, Jim Hughes, James Byron, and Yuanjiang Ni, among others. My work would not have come together if not for whiteboarding
sessions, feedback, and discussions with these fine folks. I want to explicitly thank DJ Capelis, who mentored me during
the last few years of undergrad and helped induct me into a graduate lab and give me a taste for research from which I
never recovered. I will miss you all, but don't think that you'll never see me again.

I'd additionally like to thank Robert Soulé, Avi Silberschatz, Shel Finkelstein, Andrew Quinn, Ike Nassi, Heiner Litz,
Pat Heland, Bruce Lindsay, and Ahmed Amer for their feedback and discussions. Having the ear of these brilliant people,
but more importantly listening to their insights, has been invaluable.


\squo{``I wish it need not have happened in my time,'' said Frodo.
    ``So do I,'' said Gandalf, ``and so do all who live to see such times. But that is not for them to decide. All we have
    to decide is what to do with the time that is given us.''}{\emph{The Lord of the Rings}, J.R.R. Tolkien}

Unfortunately, my time in graduate school will forever be tied to the COVID-19 pandemic. I am forever grateful to my
labmates for working hard to maintain morale. I will cherish our remote tea times. My heart goes out to all graduate
students who suffered in academia during these last few years.

Finally, all of my friends and family who stuck with me and supported me over the years. You are all amazing, and I love each and
every one of you. My parents and sister never ceased in their unending support, and I cannot thank them enough. This
dissertation is dedicated to them, from the bottom of my heart.

Thank you, all. \heart \\

{
\hfill
\emph{
    Daniel Bittman
}
}

\endgroup
